---
title: 'Lab 7, Part 2: Kriging'
author: "Sachiko Lamen"
date: "2/17/2022"
output: html_document
---

```{r setup, include= TRUE, messages = FALSE, warning = FALSE}
knitr::opts_chunk$set(echo = TRUE, messages = FALSE, warning = FALSE)
library(tidyverse)
library(here)
library(sf)
library(tmap)
library(gstat)
library(stars) # spatial temporal arrays -- will eventually take the place of the raster package
```

### Read in CA map and groundwater depth data

```{r}
### Read in counties
ca_counties_sf <- read_sf(here("data", "ca_counties", "CA_Counties_TIGER2016.shp")) %>%
  janitor::clean_names() %>%
  select(name)

### Check the projection
st_crs(ca_counties_sf)

### Read in groundwater depth data
ca_dgw_sf <- read_sf(here("data", "ca_dgw", "F2013_DBGS_Points_20150720_093252.gpkg")) %>%
  janitor::clean_names()

### Check the projection
st_crs(ca_dgw_sf)

# uh oh, they have different coordinate systems so we need to transform them to match:

ca_dgw_sf <- st_transform(ca_dgw_sf, st_crs(ca_counties_sf))

ggplot() +
  geom_sf(data = ca_counties_sf) +
  geom_sf(data = ca_dgw_sf, aes(color = dgbs))
```

```{r}
### lets do an interactive tmap
tmap_mode(mode = 'view')

tm_shape(ca_dgw_sf) +
  tm_dots('dgbs', palette = 'Blues')

## tmaptools::palette_explorer()  ### need shinyjs package!
```

From these graphs we see that there is indeed some sort of spatial pattern or correlation with groundwater depth.

### CLip to San Joaquin County
```{r}
sj_county_sf <- ca_counties_sf %>%
  filter(name == "San Joaquin")

### Now keep observations for groundwater depth in that county:
# take two spatial data sets and overlap them and take just the intersection

sj_depth_sf <- ca_dgw_sf %>%
  st_intersection(sj_county_sf)

ggplot()+
  geom_sf(data = sj_county_sf) +
  geom_sf(data = sj_depth_sf, aes(color = dgbs)) +
  theme_void() +
  scale_color_viridis_c()
```

Take points to create variogram and interpolate (guess) what ground water depth would be at other locations


 ### Check for duplicate locations
 - make sure to eliminate duplicates or variogram wont work!
 
```{r}
well_duplicates <- sj_depth_sf %>%
  janitor::get_dupes(latitude, longitude) # if lat and long are simultaneously duplicated it will let us know!

### Drop duplicated data
sj_depth_sf <- sj_depth_sf %>%
  filter(!local_well %in% well_duplicates$local_well)

```

### Create a variogram!
```{r}
## ORDINARY Kreging
sj_dgw_vgm <- gstat::variogram(dgbs ~ 1, data = sj_depth_sf)

plot(sj_dgw_vgm)
```
```{r}
### Create a variogram model by fitting a variogram to the points (above)
sj_dgw_vgm_fit <- fit.variogram(sj_dgw_vgm, 
                                model = vgm(model = 'Gau', # Gau = gaussian
                                            nugget = 20,
                                            sill = 3000,
                                            range = 32000))

plot(sj_dgw_vgm, sj_dgw_vgm_fit)

attr(sj_dgw_vgm_fit, 'SSErr') # to check the sum of squares value to determine fit we can see that the values we manually put in are better fit than the model generated values

### Based on variogram fit pass those measurements into the krigen thing 
```

### Spatial Kriging

```{r}
sj_grid <- st_bbox(sj_county_sf) %>%
  st_as_stars(dx = 1000, dy = 1000) %>%
  st_set_crs(st_crs(sj_county_sf)) %>%
  st_crop(sj_county_sf)

sj_grid

plot(sj_grid)

### perform ordinary kriging -- give it a formula (simple, ordinary, etc), and give it location (points)
sj_dgw_krige <- krige(dgbs ~ 1, sj_depth_sf, sj_grid, 
                      model = sj_dgw_vgm_fit)

plot(sj_dgw_krige)

krige_df <- as.data.frame(sj_dgw_krige) %>%
  st_as_sf(coords = c('x', 'y')) %>%
  drop_na(var1.pred)

ggplot(data = krige_df) +
  geom_sf(aes(color = var1.pred)) +
  scale_color_gradient(low = 'blue', high = 'yellow')
```

 
 
 
 
 
 
 
 
 
 
 
 
 
 
 
 
 